Quality Criteria for SKOS Vocabularies

### 2012-03-09 paper.tex
added wcc and cycles

\usepackage{verbatim}
\usepackage{url}

@@

To calculate the number of loose concepts, we let \(G = (C, SR \cup AR)\)  and apply \(deg(c)\) as a function that calculates the degree, hence the number of in- and outgoing edges of a node n in a graph g. We can then define this quality criterion as a function \(looseConcepts : G \rightarrow ℕ0\), with
\begin{definition}
looseConcepts(g) = \left\|\left\{c \in C : deg(c) = 0\right\}}\right\|
\end{definition}

>>>>

\usepackage{verbatim}
\usepackage{url}
\usepackage{amssymb}

@@

\begin{definition}
To calculate the number of loose concepts, we let \(G = (C, SR \cup AR)\)  and apply \(deg(c)\) as a function that calculates the degree, hence the number of in- and outgoing edges of a node n in a graph g. We can then define this quality criterion as a function \(looseConcepts : G \rightarrow \mathbb{N}_{0}\), with \[looseConcepts(g) = \left\|\left\{c \in C : deg(c) = 0\right\}\right\|\]
\end{definition}

\subsubsection{Weakly Connected Components} cause the vocabulary to split into separate disconnected components. This is very similar to the ``loose concepts” criterion introduced before and might be caused by incomplete data acquisition, ``forgotten" test data, outdated terms, accidental deletion of relations and the like. In a practical setting, existence of weakly connected components could render the vocabulary less suitable for operations that rely on navigating a strongly connected vocabulary structure, such as query expansion or suggestion of related terms.

\begin{definition}
To evaluate this criterion we consider a graph \(G = (AC, SRAR)\). The function \(components(g) : G \rightarrow  \mathfrak{P}(G)\) calculates all weakly connected components of a graph by replacing all directed edges by undirected edges and utilizing “Tarjan’s algorithm” \cite{Hopcroft1973}. Based on that function, we can express this criterion as a function \(wcc : G \rightarrow \mathbb{N}_{0}\), with \[wcc(g) = \left\|components(g)\right\|\]
\end{definition}

\subsubsection{Cyclic Hierarchical Relations} have been mentioned numerous times in thesaurus development literature. \cite{Soergel2002} suggests a ``check for hierarchy cycles" since they ``throw the program for a loop in the generation of a complete hierarchical structure”. However, the SKOS documentation does not state, how exactly hierarchical relationships are to be interpreted. There exist common forms like, e.g., “generic-specific”, “instance-of” or “whole-part” \cite{Hedden2011,Harpring2010,Aitchison2000} and hierarchical cycles for some of these interpretations in various domains or usage scenarios might be perfectly valid, whereas they would be treated as failures in others.

\begin{definition}
In the context of this work, we define a cycle in a SKOS vocabulary as a logical contradiction in the broader/narrower hierarchy. Let \(HB \subseteq SR\) be the set of broader hierarchical relations associating two conceptual resources with instances of \texttt{skos:broader}, \texttt{skos:broaderTransitive}, \texttt{skos:broadMatch} or subproperties thereof. \(HB\) also associates conceptual resources related with instances \texttt{skos:narrower}, \texttt{skos:narrowTransitive}, \texttt{skos:narrowMatch} or subproperties thereof in reverse order. Likewise, \(HN \subseteq SR\) can be defined as the set of narrower hierarchical relations. We consider two graphs \(GB = (C,HB)\) and \(GN=(C,HN)\) and a function \(cycleNodes:GB \cup GN \rightarrow \mathfrak{P}(C)\) identifying all nodes in a graph, that are part of a cycle. Furthermore we consider a function \(stronglyConnectedSets: GB \cup GN \rightarrow \mathfrak{P}(C)\) that calculates all strongly connected undirected subgraphs of a graph. This criterion can be defined as a function \(chr:GB \cup GN \rightarrow \mathbb{N}_{0}\) with \[chr(g)=\left\|\left\{s \in stronglyConnectedSets(g) : cn \in cycleNodes(g), cn \in s\right\}\right\|\]
\end{definition}
